---
title: "The Abyss Test: collection overview"
date: today
format:
  html:
    embed-resources: true
    toc: false
execute:
  echo: false
  warning: false
---

<!--
How long each level takes, read off what preprocess.R wrote. Run that first,
with --long, since the times an item was shown are in the long table alone:

    Rscript data/collected/preprocess.R --long
    quarto render data/collected/overview.qmd

Base R and nothing else, like preprocess.R. The page it renders is aggregates
only, but out of a handful of people an aggregate is close to a person, so it
is git-ignored beside clean/ and not to be committed.
-->

```{r}
#| label: read

clean <- "clean"
long_file <- file.path(clean, "responses.csv")
wide_file <- file.path(clean, "data.csv")
if (!file.exists(long_file)) {
  stop("No clean/responses.csv: run `Rscript data/collected/preprocess.R --long` first.")
}

long <- read.csv(long_file, stringsAsFactors = FALSE, na.strings = "NA")
wide <- read.csv(wide_file, stringsAsFactors = FALSE, na.strings = "NA")
SEP <- " | "
when <- function(x) as.POSIXct(x, format = "%Y-%m-%dT%H:%M:%OS", tz = "UTC")
```

```{r}
#| label: levels

# Which level an item was on is not a column of either table, and does not
# have to be: every level screen (`Level_<N>`) stands in the file after the
# last item of the level it showed, so the items between two screens are one
# level, and the Nth run of them is the Nth level in `levels_walked`. The
# items after the last screen are the level being answered when the file was
# written (or, in a finished run, the closing level, which has no screen).
levels_of <- function(one) {
  row <- wide[wide$participant == one, ]
  walked <- strsplit(row$levels_walked, SEP, fixed = TRUE)[[1]]
  items <- long[long$participant == one, ]
  items <- items[order(items$order), ]
  screen <- grepl("^Level_[0-9]+$", items$key)
  at <- 1 + cumsum(screen) - screen
  data.frame(
    participant = one,
    key = items$key,
    screen = screen,
    level = walked[at],
    position = at,
    onset = when(items$time_onset),
    response = when(items$time_response),
    stringsAsFactors = FALSE
  )
}
items <- do.call(rbind, lapply(unique(wide$participant), levels_of))

# A level counts once it was finished: every item of it answered, which is
# what `QC_<key>_TimeFinished` says. A level left partway would put the time
# somebody gave up into the figure for how long the level takes.
finished <- function(one, key) {
  column <- paste0("QC_", key, "_TimeFinished")
  column %in% names(wide) && !is.na(wide[wide$participant == one, column])
}

# From the first item of the level being put on screen to the last being
# answered, briefings included and the results screen after it not: wall time,
# so a pause in the middle is in it. An item shown again has its onset
# re-stamped later, which the earliest onset of the level mostly survives.
spans <- do.call(rbind, lapply(split(items[!items$screen, ], list(items$participant[!items$screen], items$level[!items$screen]), drop = TRUE), function(one) {
  data.frame(
    participant = one$participant[1],
    level = one$level[1],
    position = one$position[1],
    minutes = as.numeric(difftime(max(one$response, na.rm = TRUE), min(one$onset, na.rm = TRUE), units = "mins")),
    finished = finished(one$participant[1], one$level[1]),
    stringsAsFactors = FALSE
  )
}))
done <- spans[spans$finished, ]

# The results screen of a level, from its being uncovered to the way on being
# pressed: the one place the file measures how long results were read.
screens <- items[items$screen & !is.na(items$onset) & !is.na(items$response), ]
screens$minutes <- as.numeric(difftime(screens$response, screens$onset, units = "mins"))
```

```{r}
#| label: summary

# Answering is in minutes and the results screen in seconds: the one runs to
# half an hour a level, the other to a minute or two, and neither reads well
# in the other's unit.
summarise <- function(frame, unit = "minutes") {
  if (!nrow(frame)) return(data.frame(level = character(), n = integer(), median = numeric(), min = numeric(), max = numeric()))
  do.call(rbind, lapply(split(frame, frame$level), function(one) {
    time <- one[[unit]]
    data.frame(level = one$level[1], n = nrow(one), median = median(time), min = min(time), max = max(time))
  }))
}

# Levels are listed in the order they were met on average, since the order is
# drawn and partly chosen and no one person's order is the order.
table <- summarise(done)
where <- tapply(done$position, done$level, mean)
table <- table[order(where[table$level]), ]
rownames(table) <- NULL
screens$seconds <- screens$minutes * 60
reading <- summarise(screens, "seconds")
reading <- reading[match(table$level, reading$level), ]
reading$level <- table$level
reading$n[is.na(reading$n)] <- 0L

# A level finished whose screen was never left: the way on is staged only when
# it is pressed, so a finished level with no screen in the file is somebody who
# closed the tab on its results (or is still reading them). Only a run that
# stopped can do that; a finished run has left every screen.
left <- paste(screens$participant, screens$level)
stopped_on <- done[!paste(done$participant, done$level) %in% left, ]
reading$closed <- as.integer(table(factor(stopped_on$level, levels = table$level)))
```

`r nrow(wide)` run(s), `r sum(wide$completed)` finished and `r sum(!wide$completed)` stopped partway; `r nrow(done)` level(s) finished between them, out of `r nrow(spans)` begun. Test and synthetic runs are left out unless `preprocess.R` was run with `--include-test`.

## How long a level takes

Two times for each level, each on its own scale. **Answering**, in minutes, runs from the first item of the level appearing to its last being answered, over the runs that finished that level; a level begun and left is not counted. **Results**, in seconds, runs from the level's results being uncovered to the way on being pressed, over those who pressed it.

```{r}
#| label: plot
#| fig-width: 8
#| fig-height: !expr max(2.4, 0.55 * nrow(table) + 1.4)

ANSWERING <- c(line = "#a58bff", dot = "#7c5cff80", median = "#5b3fd6")
RESULTS <- c(line = "#e8c77a", dot = "#d9a44180", median = "#b07d1c")

# One panel a measure, side by side on a shared row a level, each on an axis
# of its own: a line from the fastest run to the slowest, a dot a run and the
# median ringed. Only the left panel names the levels.
panel <- function(summary, runs, unit, title, colours, named) {
  y <- rev(seq_len(nrow(summary)))
  has <- summary$n > 0
  top <- if (any(has)) max(summary$max[has]) else 1
  par(mar = c(4, if (named) 9 else 1.5, 2, 1.5))
  plot(NULL, xlim = c(0, top * 1.1), ylim = c(0.5, nrow(summary) + 0.5),
       yaxt = "n", xlab = tools::toTitleCase(unit), ylab = "", bty = "n")
  title(main = title, font.main = 1, cex.main = 1, adj = 0)
  if (named) axis(2, at = y, labels = summary$level, las = 1, tick = FALSE)
  abline(h = y, col = "grey93")
  segments(summary$min[has], y[has], summary$max[has], y[has], lwd = 3, col = colours["line"])
  points(runs[[unit]], y[match(runs$level, summary$level)], pch = 16, cex = 0.7, col = colours["dot"])
  points(summary$median[has], y[has], pch = 21, bg = colours["median"], col = "white", cex = 1.6)
}

if (nrow(table)) {
  # The left panel is wider by the names it carries, so the two plots
  # themselves come out the same width.
  op <- par(no.readonly = TRUE)
  layout(matrix(1:2, 1), widths = c(1.45, 1))
  panel(table, done, "minutes", "Answering", ANSWERING, TRUE)
  panel(reading, screens[screens$level %in% table$level, ], "seconds", "Results screen", RESULTS, FALSE)
  par(op)
} else {
  cat("No level has been finished yet.")
}
```

```{r}
#| label: table

span <- function(summary, digits = 1) {
  ifelse(summary$n > 0, sprintf("%.*f to %.*f", digits, summary$min, digits, summary$max), "")
}
middle <- function(summary, digits = 1) ifelse(summary$n > 0, sprintf("%.*f", digits, summary$median), "")
shown <- data.frame(
  Level = table$level,
  `Answering: n` = table$n,
  `median (min)` = middle(table),
  `range (min)` = span(table),
  `Results: n` = reading$n,
  `median (s)` = middle(reading, 0),
  `range (s)` = span(reading, 0),
  `Closed on results` = reading$closed,
  check.names = FALSE
)
knitr::kable(shown, align = "lrrrrrrr")
```

*Closed on results* counts the runs that finished the level and never pressed the way on from its results, which is where a run stopped partway most often ends; they are not in the results times, having no end to measure.
